---
title: "PD2 - NSE"
author: "Jadwiga Słowik"
date: "November 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("nse_funs.R")
source("utils.R")
library(dplyr)
```

## Aim
The aim of this task was to implement wrapper functions using *NSE* (*Non standard evaluation*)
and then test them in order to measure their performance.

## Implemented functions
Four versions of wrapper-functions were implemented. In some of them I used the *base* package for *NSE*
and in the other *rlang* package was used, which provides *quasiquotation*.

The implementation of aforementioned *NSE* functions are provided in a separate file - `nse_funs.R`.

## Tests
Tests were splitted into two groups: small and big.

### Small tests

#### min
The test case has the following form:

```{r}
data <- list(a=c(1,2,3), b=c(1,2,3) * 10)
fun <- min
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

To check the result:
```{r}
c(nse1(data, b), nse2(data, b), nse3(data, b), nse4(data, b), min(data$b))
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b)},
    nse2={nse2(data, b)},
    nse3={nse3(data, b)},
    nse4={nse4(data, b)},
    raw={min(data$b)},
    unit="ms",
    times=10
  )
)
```

#### max
The test case has the following form:

```{r}
data <- list(a=c(1,2,3), b=c(1,2,3) * 10)
fun <- max
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

To check the result:
```{r}
c(nse1(data, b), nse2(data, b), nse3(data, b), nse4(data, b), max(data$b))
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b)},
    nse2={nse2(data, b)},
    nse3={nse3(data, b)},
    nse4={nse4(data, b)},
    raw={max(data$b)},
    unit="ms",
    times=10
  )
)
```

#### mean
The test case has the following form:

```{r}
data <- list(a=c(1,2,3), b=c(1,2,3) * 10)
fun <- mean
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

To check the result:
```{r}
c(nse1(data, b), nse2(data, b), nse3(data, b), nse4(data, b), mean(data$b))
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b)},
    nse2={nse2(data, b)},
    nse3={nse3(data, b)},
    nse4={nse4(data, b)},
    raw={mean(data$b)},
    unit="ms",
    times=10
  )
)
```

#### unlist

##### use.names=TRUE

The test case has the following form:

```{r}
data <- list(a=list(a=1,b=2,c=3), b=list(a=10,b=20,c=30))
fun <- unlist
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

To check the result:
```{r}
list(nse1(data, b, use.names=TRUE), nse2(data, b, use.names=TRUE), nse3(data, b, use.names=TRUE), nse4(data, b, use.names=TRUE), unlist(data$b, use.names=TRUE))
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b, use.names=TRUE)},
    nse2={nse2(data, b, use.names=TRUE)},
    nse3={nse3(data, b, use.names=TRUE)},
    nse4={nse4(data, b, use.names=TRUE)},
    raw={unlist(data$b, use.names = TRUE)},
    unit="ms",
    times=10
  )
)
```

#### unlist

##### use.names=TRUE

The test case has the following form:

```{r}
data <- list(a=list(a=1,b=2,c=3), b=list(a=10,b=20,c=30))
fun <- unlist
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

To check the result:
```{r}
list(nse1(data, b, use.names=FALSE), nse2(data, b, use.names=FALSE), nse3(data, b, use.names=FALSE), nse4(data, b, use.names=FALSE), unlist(data$b, use.names=FALSE))
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b, use.names=FALSE)},
    nse2={nse2(data, b, use.names=FALSE)},
    nse3={nse3(data, b, use.names=FALSE)},
    nse4={nse4(data, b, use.names=FALSE)},
    raw={unlist(data$b, use.names=FALSE)},
    unit="ms",
    times=10
  )
)
```


#### lm

The test case has the following form:

```{r}
data <- data.frame(a=1:10, b=1:10)
fun <- lm
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

To check the result:
```{r}
list(nse1(data, a~b)$coefficients, nse2(data, a~b)$coefficients, nse3(data, a~b)$coefficients, nse4(data, a~b)$coefficients, lm(a~b, data)$coefficients)
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, a~b)},
    nse2={nse2(data, a~b)},
    nse3={nse3(data, a~b)},
    nse4={nse4(data, a~b)},
    raw={lm(a~b, data)},
    unit="ms",
    times=10
  )
)
```

### Larger tests
Due to readablility for this kind of tests printing outputs is omitted .

### min
The test case has the following form:

```{r}
n <- 100000
data <- data.frame(a=1:n, b=1:n)
fun <- min
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, a)},
    nse2={nse2(data, a)},
    nse3={nse3(data, a)},
    nse4={nse4(data, a)},
    raw={min(data$a)},
    unit="ms",
    times=10
  )
)
```

### max
The test case has the following form:

```{r}
n <- 100000
data <- data.frame(a=1:n, b=1:n)
fun <- max
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, a)},
    nse2={nse2(data, a)},
    nse3={nse3(data, a)},
    nse4={nse4(data, a)},
    raw={max(data$a)},
    unit="ms",
    times=10
  )
)
```

#### unlist

##### use.names=TRUE

The test case has the following form:

```{r}
n <- 100000
data <- data.frame(a=1:n, b=1:n)
fun <- unlist
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b, use.names=TRUE)},
    nse2={nse2(data, b, use.names=TRUE)},
    nse3={nse3(data, b, use.names=TRUE)},
    nse4={nse4(data, b, use.names=TRUE)},
    raw={unlist(data$b, use.names = TRUE)},
    unit="ms",
    times=10
  )
)
```

#### unlist

##### use.names=TRUE

The test case has the following form:

```{r}
n <- 100000
l <- as.list(1:n)
names(l) = 1:n
data <- list(b=l)
fun <- unlist
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, b, use.names=FALSE)},
    nse2={nse2(data, b, use.names=FALSE)},
    nse3={nse3(data, b, use.names=FALSE)},
    nse4={nse4(data, b, use.names=FALSE)},
    raw={unlist(data$b, use.names=FALSE)},
    unit="ms",
    times=10
  )
)
```


#### lm

The test case has the following form:

```{r}
n <- 1000000
data <- data.frame(a=1:n, b=1:n)
fun <- lm
nse1 <- my_NSE1(fun)
nse2 <- my_NSE2(fun)
nse3 <- my_NSE3(fun)
nse4 <- my_NSE4(fun)
```

The benchmark is the following:
```{r}
visualize(
  microbenchmark::microbenchmark(
    nse1={nse1(data, a~b)},
    nse2={nse2(data, a~b)},
    nse3={nse3(data, a~b)},
    nse4={nse4(data, a~b)},
    raw={lm(a~b, data)},
    unit="ms",
    times=10
  )
)
```

## Summary
In all cases the *raw* version of a functions turned out to be the most efficient.
Moreover, in most cases the difference in performance is very significant.
Additionally, it is worth pointing out that in all executed tests the provided functions performed similarly
taking into account the performance rank. Namely, the rank is: *raw*, *nse4*, *nse2*, *nse1*, *nse3*.
Usually, the difference between *nse3* and *nse1* is larger than the other.

However, a reader should notice that in the last test (larger test for *lm*) the differences in performance are negligible - all perform similarly.
