---
title: "graphs"
output: html_document
---
Graphs generated with different libraries. 

ggplot_pointchart

```{r}
library(ggplot2)
library(lattice)

data = data.frame(x = runif(100), y = runif(100))

ggplot(data = data, aes(x = data$x, y = data$y)) +
    geom_point()
```
ggplot_linechart
```{r}
  ggplot(data = data, aes(x = data$x, y = data$y)) +
    geom_line()
```
ggplot_barchart
```{r}
  ggplot(data = data, aes(
    x = data$x, y = data$y, col = data$x
  )) +
    geom_bar(stat = "identity")
```
ggplot_boxchart
```{r}
  ggplot(data = data, aes(
    x = data$x, y = data$y, col = data$x
  )) + geom_boxplot()
```
ggplot_densitychart
```{r}
  ggplot(data = data, aes(x = data$x)) +
    geom_histogram(binwidth = 1)
```
ggplot_nochart
```{r}
  ggplot()
```
lattice_pointchart
```{r}
  xyplot(y ~ x, data)
```
lattice_linechart
```{r}
  xyplot(y ~ x, data, type = c("smooth"))
```
lattice_barchart
```{r}
  barchart(y ~ x, data)
```
lattice_boxchart
```{r}
  bwplot(y ~ x, data)
```
lattice_densitychart
```{r}
  histogram(data$x)
```
lattice_nochart
```{r}
  xyplot(0 ~ 0, data = data)
```
graphics_pointchart
```{r}
  plot(data$x, data$y, type = "p")
```
graphics_linechart
```{r}
  plot(data$x, data$y, type = "l")
```
graphics_barchart
```{r}
  plot(data$x, data$y, type = "h")
```
graphics_boxchart
```{r}
  boxplot(data$x, data$y)
```
graphics_densitychart
```{r}
  hist(data$x)
```
graphics_nochart
```{r}
  plot(data$x, data$y, type = "n")
```