---
title: "Praca domowa 1"
author: "Sebastian Sowik"
date: "October 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(wesanderson)
library(tidyr)
library(lattice)
library(latticeExtra)
library(microbenchmark)

data(msleep, package="ggplot2")

orderedNames <- msleep %>% arrange(sleep_total) %>% mutate(name=factor(name,name)) %>% filter(conservation=="domesticated") %>% pull(name) %>% as.factor()
barplotData <- msleep %>% filter(conservation=="domesticated") %>% 
  gather("state", "time", sleep_total, awake)  %>% mutate(
    state=factor(ifelse(state=="sleep_total", "asleep", "awake"),c("awake","asleep")),
    name=factor(name,orderedNames)
  ) 

ggplot_boxplot <- function () {
  ggplot(data=na.omit(msleep), aes(x=vore, y=sleep_total)) +
    geom_boxplot(colour="black") +
    geom_point(color="blue") +
    theme_bw() +
    labs(x="Vore", y="Hours of sleep per day", title="Distribution of time of sleep per vore")
}

lattice_boxplot <- function () {
  bwplot(
    sleep_total~vore,
    msleep,
    xlab="Vore",
    ylab="Hours of sleep per day",
    main="Distribution of time of sleep per vore",
    par.settings=list(
      box.umbrella = list(col = "black"), 
      box.rectangle=list(col= rep(c("black", "black"),2)))
    ) +
  dotplot(sleep_total~vore, msleep, col="blue")
}

graphics_boxplot <- function () {
  boxplot(sleep_total~vore,
          msleep,
          xlab="Vore",
          ylab="Hours of sleep per day",
          main="Distribution of time of sleep per vore"
          )
  stripchart(sleep_total~vore, vertical = TRUE, data = msleep, 
             method = "jitter", add = TRUE, pch = 20, col = 'blue')
}


ggplot_barplot <- function () {
  barplotData %>% 
    select(name, time, state) %>% 
    ggplot(aes(x=name, y=time, fill=state)) +
    geom_bar(stat="identity") +
    labs(x="Specie",
         y="Duration of state",
         title="Sleep awake ratio for domesticated animals") +
    theme_minimal()
}


lattice_barplot <- function () {
barchart(time ~ name,
         groups=factor(barplotData[["state"]], c("asleep", "awake")),
         data=barplotData,
         stack=TRUE,
         xlab="Specie",
         ylab="Duration of state",
         auto.key=list(space="right", title="State"),
         main="Sleep awake ratio for domesticated animals")
}
graphicsData <- barplotData %>% select(name, state, time) %>% spread(name, time) %>% select(-state) %>% arrange(Horse) %>%as.matrix.data.frame()
rownames(graphicsData) <- c("awake", "asleep")

graphics_barplot <- function() {
  barplot(graphicsData,
        legend=c("awake", "asleep"),
         xlab="Specie",
         ylab="Duration of state",
         main="Sleep awake ratio for domesticated animals")
}

sizeData = read.csv("./MemoryForNestedLists.csv")

grid_heatmaps_ggplot <- function() {
  sizeData %>% ggplot(aes(x=Outermost, y=Innermost, fill=Size)) +
    geom_raster() +
    facet_grid(Nested~Nestings, labeller = label_both) +
    scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) +
    labs(title="Memory usage in nested list structure",
         fill="Size in bytes")
}

grid_heatmaps_lattice <- function() {
  levelplot(Size~Innermost*Outermost|Nestings*Nested,
          sizeData,
          auto.key=list(space="right",title="Size in bytes"))
}

contourDataList = list()
for(i in 1:5) {
  for(j in 1:5) {
    data <- sizeData %>% filter(Nested==i & Nestings==j) %>% 
            select(-Nested, -Nestings, -X) %>% spread(Outermost, Size) %>% data.matrix()
    contourDataList <- append(contourDataList, list(data))
  }
}

source("./Filled.contour3.R")
grid_heatmaps_graphics <- function() {
  par(mfrow=c(5,5), mar=rep(1,4)) 
  for(i in 1:25) {
    filled.contour3(z=contourDataList[[i]] )
  }
}
sizeData %>% mutate(Elements=Outermost*Innermost*Nested, SizeByElems=(Size/(Elements))) -> processedData

scattered_facet_ggplot <- function() {
  processedData %>% ggplot(aes(x=Elements, y=Size, col=factor(Innermost))) +
    geom_point() +
    facet_wrap(~Nestings, labeller = label_both) +
    labs(y="Size in bytes",
         col="Nr of elements\nin the most nested lists",
         title="Size of nested lists"
         )
}


palette <- rainbow(50)
scattered_facet_lattice <- function() {
  xyplot(
    Size~Elements|Nestings,
    processedData,
    groups=Innermost,
    col=palette,
    pch=19,
    key = list(space="right",
               columns=3,
               title="Nr of elements \nin the most nested lists",
               points=list(col=palette, lty=c(3,2), lwd=3),
               text=list(as.character(1:50))
              ),
    ylab="Size in bytes"
    )
}


xygraphicsdata <- processedData %>% filter(Nestings==1) %>% select(Size, Elements)
graphicsScatterData <- list()
for(i in 1:5) {
  graphicsdata <- processedData %>% filter(Nestings==i) %>% select(Size, Elements)
  graphicsScatterData <- append(graphicsScatterData, list(graphicsdata))
}

scattered_facet_graphics <- function() {
  par(mfrow=c(2,3))
  for (i in 1:5){
    plot(Size~Elements, graphicsScatterData[[i]], col=palette)
  }
}



groupedSizeData <- processedData %>% group_by(Nested, Nestings) %>%
  summarise(mSizeByElems=mean(SizeByElems))

heatmap_ggplot <- function() {
  groupedSizeData %>% 
    ggplot(aes(x=Nestings, y=Nested, fill=mSizeByElems)) +
    geom_tile() +
    scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) +
    labs(fill="Mean size\nper element\n[bytes/elem]",
         title="Memory eficiency for lists with different levels of nesting")
}

heatmap_lattice <- function() {
  levelplot(mSizeByElems~Nestings*Nested, groupedSizeData,
            main="Memory eficiency for lists with different levels of nesting",
            legend=list(top=list(fun=grid::textGrob("Mean size\nper element\n[bytes/elem]", y=0, x=1.06)))
            )
}
graphicsHeatmapData <- groupedSizeData %>% spread(Nested, mSizeByElems) %>% select(-Nestings) %>% 
  data.matrix()
heatmap_graphics <- function() {
  filled.contour(graphicsHeatmapData,
                 plot.title=title(main="Memory eficiency for lists with different levels of nesting"),
                 key.title=title(main=list("Mean size\nper element\n[bytes/elem]", cex=0.8))
  )
}
```

## Tested plots

```{r ggplot_boxplot}
ggplot_boxplot()
```

```{r lattice_boxplot}
lattice_boxplot()
```

```{r graphics_boxplot}
graphics_boxplot()
```

```{r ggplot_barplot}
ggplot_barplot()
```

```{r lattice_barplot}
lattice_barplot()
```

```{r graphics_barplot}
graphics_barplot()
```

```{r grid_heatmaps_ggplot}
grid_heatmaps_ggplot()
```

```{r grid_heatmaps_lattice}
grid_heatmaps_lattice()
```

```{r grid_heatmaps_graphics}
grid_heatmaps_graphics()
```

```{r scattered_facet_ggplot}
scattered_facet_ggplot()
```

```{r scattered_facet_lattice}
scattered_facet_lattice()
```

```{r scattered_facet_graphics}
scattered_facet_graphics()
```

```{r heatmap_ggplot}
heatmap_ggplot()
```

```{r heatmap_lattice}
heatmap_lattice()
```

```{r heatmap_graphics}
heatmap_graphics()
```

## Benchmark results


```{r, fig.show = 'hide'}

microbenchmark(
  ggplot_boxplot=ggplot_boxplot,
  lattice_boxplot=lattice_boxplot,
  graphics_boxplot=graphics_boxplot
)
microbenchmark(
  ggplot_barplot=ggplot_barplot,
  lattice_barplot=lattice_barplot,
  graphics_barplot=graphics_barplot
)
microbenchmark(
  grid_heatmaps_ggplot=grid_heatmaps_ggplot,
  grid_heatmaps_lattice=grid_heatmaps_lattice,
  grid_heatmaps_graphics=grid_heatmaps_graphics
)
microbenchmark(
  scattered_facet_ggplot=scattered_facet_ggplot,
  scattered_facet_lattice=scattered_facet_lattice,
  scattered_facet_graphics=scattered_facet_graphics
)

microbenchmark(
  heatmap_ggplot=heatmap_ggplot,
  heatmap_lattice=heatmap_lattice,
  heatmap_graphics=heatmap_graphics
)


```

## Conclusions

Execution time of tested plots is generaly similar, there is no clear winner. 
Graphics requires more preprocessing of data, contour plots does not support facets. 
Lattice is more convenient thanks to formulas but plot customization is cumbersome.
ggplot2 is the most convenient library with intuitive layer system and flexible customization.
