---
title: "Advanced R"
author: "Tymoteusz Makowski"
date: '2019-10-10'
subtitle: Home assignment (PD1)
---


# Data

Data represents information about departs from Świnoujście port in Poland ([Source](https://dane.gov.pl/)). 
```{r, include = FALSE}
n_rep <- 50

knitr::opts_chunk$set(comment = NA, echo = FALSE, message = FALSE,
                      warning = FALSE, error = FALSE)

library("ggplot2")
library("graphics")
library("data.table")

df <- setNames(
    fread("~/dane.csv", encoding = "Latin-1"),
    c("DepartCode", "Date", "Hour", "DestCode", "ShipName", "Flag", "id",
      "Signal", "Volume", "Length", "Agent", "DockingPlace")
)

df_agg_month <- df[,
    .(
       Flag,
       DateAgg = strftime(as.Date(Date), "%Y-%m")
    )
][
, .(nDeparts = .N), by = .(Flag, DateAgg)
][
   , .(
       Flag = ifelse(nDeparts >= 10, Flag, "Other"),
       DateAgg,
       nDeparts
   ) 
][
   , .(nDeparts = sum(nDeparts)), by = .(Flag, DateAgg)
]
```


# Plot 1. - Line plot

## Plots

### `ggplot2`
```{r}
plot1_ggplot2 <- function() {
    ggplot(df_agg_month, aes(x = DateAgg, y = nDeparts, color = Flag,
                             group = Flag)) +
        geom_line() +
        theme_classic() +
        labs(x = "Date", y = "Number of departs",
             title = "Departs from Świnoujście port") +
        scale_x_discrete(breaks = paste0(2008:2019, "-01")) +
        theme(
            axis.text.x = element_text(angle = 45, hjust = 1)
        )
}

plot1_ggplot2()
```


### `graphics`
```{r, fig.width = 7.3}
flags <- sort(unique(df_agg_month[["Flag"]]))
flag_cols <- setNames(
    scales::hue_pal()(length(flags)),
    flags
)

date_new <- as.Date(paste0(df_agg_month$DateAgg, "-01"))
x_lim <- c(min(date_new), max(date_new))
y_lim <- c(min(df_agg_month$nDeparts), max(df_agg_month$nDeparts))

plot_single <- function(flag, plot_func = lines, df = df_agg_month,
                        flags_colors = flag_cols, ...) {
    df_filtered <- df[Flag == flag]
    y_vals <- df_filtered$nDeparts
    x_vals <- as.Date(paste0(df_filtered$DateAgg, "-01"))
    
    plot_func(y_vals ~ x_vals, xlim = x_lim, ylim = y_lim,
              col = flags_colors[[flag]], ...)
}

plot1_graphics <- function() {
    par(mar =  c(5, 4, 4, 13) + 0.1)
    plot_single(flags[1], plot, type = "l", xlab = "Date",
                main = "Departs from Świnoujście port",
                ylab = "Number of departs")
    for (f in flags[-1]) {
        plot_single(f)
    }
    legend(x = "topright", legend = names(flag_cols), col = flag_cols, lty = 1,
           inset = c(-0.65, 0), title = "Flag", xpd = TRUE)
}

plot1_graphics()
```

### `lattice`
```{r}
plot1_lattice <- function() {
    lattice::xyplot(nDeparts ~ as.Date(paste0(DateAgg, "-01")), df_agg_month,
                    groups = Flag, type = "l",
                    xlab = "Date", ylab = "Number of departs",
                    auto.key = TRUE
    )
}

plot1_lattice()
```

## Comparison

```{r, fig.show = 'hide'}
plot1_mb <- microbenchmark::microbenchmark(
    ggplot2 = plot1_ggplot2(),
    graphics = plot1_graphics(),
    lattice = plot1_lattice(),
    times = n_rep
)
```

```{r}
autoplot(plot1_mb) + theme_classic()
```



# Plot 2. - Colored point plot

## Plots

### `ggplot2`
```{r}
plot2_ggplot2 <- function() {
    ggplot(df_agg_month, aes(x = DateAgg, y = nDeparts, color = Flag)) +
        geom_point() +
        theme_classic() +
            labs(x = "Date", y = "Number of departs",
                 title = "Departs from Świnoujście port") +
        scale_x_discrete(breaks = paste0(2008:2019, "-01")) +
            theme(
                axis.text.x = element_text(angle = 45, hjust = 1)
            )
}

plot2_ggplot2()
```

### `graphics`

```{r}
plot2_graphics <- function() {
    par(mar =  c(5, 4, 4, 13) + 0.1)
    plot_single(flags[1], plot, type = "p", xlab = "Date",
                main = "Departs from Świnoujście port",
                ylab = "Number of departs")
    for (f in flags[-1]) {
        plot_single(f, points, pch = 16)
    }
    legend(x = "topright", legend = names(flag_cols), col = flag_cols, lty = 1,
           inset = c(-0.65, 0), title = "Flag", xpd = TRUE)
}

plot2_graphics()
```

### `lattice`
```{r}
plot2_lattice <- function() {
    lattice::dotplot(nDeparts ~ DateAgg, df_agg_month, col = flag_cols,
                     main = "Departs from Świnoujście port",
                     xlab = "Date", ylab = "Number of departs",
                     key = list(
                       space = "right",
                       lines = list(col = flag_cols),
                       text = list(names(flag_cols))
                     ),
                     scales = list(
                       x = list(
                         labels = sapply(2008:2019, function(x) c(x, rep("", 11)))
                       )
                     )
    )
}

plot2_lattice()
```

## Comparison
```{r, fig.show = 'hide'}
plot2_mb <- microbenchmark::microbenchmark(
    ggplot2 = plot2_ggplot2(),
    graphics = plot2_graphics(),
    lattice = plot2_lattice(),
    times = n_rep
)
```

```{r}
autoplot(plot2_mb) + theme_classic()
```


# Plot 3. - Bar plot

## Plots
```{r}
agents <- unlist(df[, .(n = .N), by = .(Agent)][order(-n)][n > 500, .(Agent)])
```

### `ggplot2`
```{r}
plot3_ggplot2 <- function() {
    ggplot(df[Agent %in% agents], aes(x = Agent)) +
        geom_bar() +
        scale_x_discrete(limits = agents) +
        theme_classic() +
            labs(x = "Agent", y = "Number of departs",
                 title = "Top agents (with >500 departs)") +
        
            theme(
                axis.text.x = element_text(angle = 45, hjust = 1)
            )
}

plot3_ggplot2()
```

### `graphics`
```{r}
plot3_graphics <- function() {
    par(mar = c(19.5, 3, 0.5, 3) + 0.1)
    barplot(sort(df[Agent %in% agents, .(n = .N), by = .(Agent)]$n,
                 decreasing = TRUE), names.arg = agents, las = 2)
}

plot3_graphics()
```

### `lattice`
```{r}
plot3_lattice <- function() {
    lattice::barchart(Count ~ reorder(Agent, -Count),
                      df[Agent %in% agents][, .(Count = .N), by = .(Agent)])
}

plot3_lattice()
```


## Comparison
```{r, fig.show = 'hide'}
plot3_mb <- microbenchmark::microbenchmark(
    ggplot2 = plot3_ggplot2(),
    graphics = plot3_graphics(),
    lattice = plot3_lattice(),
    times = n_rep
)
```

```{r}
autoplot(plot3_mb) + theme_classic()
```

# Plot 4. - Point plot (many points)
## Plots
### `ggplot2`

```{r, fig.height = 15}
plot4_ggplot2 <- function() {
    ggplot(df, aes(x = Volume, y = Agent)) +
        geom_point() +
        theme_classic() +
        ggtitle("Ship volume by agent")
}

plot4_ggplot2()
```

### `graphics`
```{r, fig.height = 30}
plot4_graphics <- function() {
    par(mar = c(4, 10, 1, 1))
    plot(df$Volume, 1:nrow(df), type = "p")
    axis(2, at = 1:nrow(df), labels = df$Agent, las = 1, lwd = 0.1)
}

plot4_graphics()
```

### `lattice`
```{r, fig.height = 30}
plot4_lattice <- function() {
    lattice::dotplot(1:nrow(df) ~ Volume, df)
}

plot4_lattice()
```

## Comparison
```{r, fig.show = 'hide'}
plot4_mb <- microbenchmark::microbenchmark(
    ggplot2 = plot4_ggplot2(),
    graphics = plot4_graphics(),
    lattice = plot4_lattice(),
    times = n_rep
)
```

```{r}
autoplot(plot4_mb) + theme_classic()
```

# Plot 5. - Histogram
## Plots
### `ggplot2`

```{r}
plot5_ggplot2 <- function () {
    ggplot(df, aes(x = Volume)) +
        geom_histogram() +
        theme_classic() +
        ggtitle("Ship volume histogram")
}

plot5_ggplot2()
```

### `graphics`

```{r}
plot5_graphics <- function() {
    hist(df$Volume, main = "Ship volume histogram")
}

plot5_graphics()
```

### `lattice`
```{r}
plot5_lattice <- function() {
    lattice::histogram( ~ Volume, data = df, main = "Ship volume histogram")
}

plot5_lattice()
```


## Comparison
```{r, fig.show = 'hide'}
plot5_mb <- microbenchmark::microbenchmark(
    ggplot2 = plot5_ggplot2(),
    graphics = plot5_graphics(),
    lattice = plot5_lattice(),
    times = n_rep
)
```

```{r}
autoplot(plot5_mb) + theme_classic()
```

# Conclusion
`lattice` package is a fastest when it comes to plotting unless there is
a large dataset involved where it is the slowest package.
`ggplot2` excels at plotting large datasets where it is way more faster than the
other two.
`graphics` package generally is the slowest one overall.

If I were to choose a package for plotting it would come down to choice between
`lattice` and `ggplot2`. However I would pick `ggplot2` over `lattice` due to
more intuitive interface (simple things with `lattice` get tedious at times)
and the speed of plotting being comparable (or faster with huge datasets).