---
title: "Praca domowa 1"
author: "Jakub Pingielski"
date: "16/10/2019"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show='hold', fig.align = 'center',
                      autodep=TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(lattice)
library(tidyverse)
library(microbenchmark)
```

## 1.1 Boxplots
### Using data frame `rock` - containing measurements on 48 rock samples from a petroleum reservoir.
 
```{r Boxplots}
ggplot_boxplot <- function() {
    ggplot(data = rock, aes(x = factor(perm), y = area)) + 
        geom_boxplot() +
        xlab("permeability [milli-Darcies]") + 
        ylab("Area of pores space")
}

ggplot_boxplot()

lattice_boxplot <- function() {
    bwplot(area ~ factor(perm), rock,
           xlab = "permeability [milli-Darcies]", 
           ylab = "Area of pores space")
} 

lattice_boxplot()

graphics_boxplot <- function() {
    boxplot(area ~ factor(perm), rock,
            xlab = "permeability [milli-Darcies]", 
            ylab = "Area of pores space")
}

graphics_boxplot()
```

## 1.2 Scatterplots
### Using data frame `trees` - containing measurements of the diameter, height and volume of timber in 31 felled black cherry trees

```{r Scatterplots}
ggplot_scatter <- function() {
    ggplot(data = trees, aes(x = Girth, y = Height)) + 
        geom_point()
}

ggplot_scatter()

lattice_scatter <- function() {
    xyplot(Height ~ Girth, trees)
}

lattice_scatter()

graphics_scatter <- function() {
    plot(Height ~ Girth, trees)
}

graphics_scatter()
```


## 1.3 Time series plots
### Using data frame `LakeHuron` - containing annual measurements of the level, in feet, of Lake Huron 1875–1972.

```{r Time series}
# Adding year column to dataset about LakeHuron levels
LakeHuron <- data.frame(LakeHuron, 1875:1972)
names(LakeHuron) <- c("Level", "Year")

ggplot_ts <- function() {
    ggplot(data = LakeHuron, aes(Year, Level)) + 
        geom_line()
}

ggplot_ts()

lattice_ts <- function() {
    xyplot(Level ~ Year, LakeHuron, type="l")
}

lattice_ts()

graphics_ts <- function() {
    plot(LakeHuron$Year, LakeHuron$Level,  type="l",
         xlab = "Year", 
         ylab = "Level")
}

graphics_ts()
```

## 1.4 Histograms
### Using data frame `Nile` - containing measurements of the annual flow of the river Nile at Aswan, 1871–1970, in 10^8 m^3.

```{r Histograms}
Nile <- as.data.frame(Nile)

ggplot_histogram <- function() {
    ggplot(Nile, aes(x)) + 
        geom_histogram() +
        xlab("Annual flow [10^8 m^3]") + 
        ylab("Count")
}

ggplot_histogram()

lattice_histogram <- function() {
    histogram(~ x, Nile, type = "count",
              xlab = "Annual flow [10^8 m^3]", 
              ylab = "Count")
}

lattice_histogram()

graphics_histogram <- function() {
    hist(Nile$x,
         xlab = "Annual flow [10^8 m^3]", 
         ylab = "Count",
         main=NA)  
}

graphics_histogram()
```

## 1.5 Density plots
### Using data frame `PlantGrowth` - containing measurements of dried weight of plants obtained under a control and two different treatment conditions.
```{r Density plots}
ggplot_density <- function() {
    ggplot(PlantGrowth, aes(weight, ..count.., color = fct_rev(group))) +
        geom_density() + 
        xlab("Weight") + 
        ylab("Density") +
        labs(color = "Group")
}

ggplot_density()

lattice_density <- function() {
    densityplot(~weight,data=PlantGrowth,
                groups=group,
                plot.points=FALSE)
}

lattice_density()

graphics_density <- function() {
    plot(density(PlantGrowth$weigh[PlantGrowth$group == 'trt2']), 
         col='green', xlim=c(2,7), xlab="Density", main=NA)
    
    lines(density(PlantGrowth$weight[PlantGrowth$group == 'trt1']),col="red")
    
    lines(density(PlantGrowth$weight[PlantGrowth$group == 'ctrl']),col="blue")
}   
graphics_density()
```

## 2 Benchmarks

```{r Benchmarks, fig.show='hide'}
microbenchmark(
    ggplot_boxplot(),
    lattice_boxplot(),
    graphics_boxplot(),
    ggplot_scatter(),
    lattice_scatter(),
    graphics_scatter(),
    ggplot_ts(),
    lattice_ts(),
    graphics_ts(),
    ggplot_histogram(),
    lattice_histogram(),
    graphics_histogram(),
    ggplot_density(),
    lattice_density(),
    graphics_density(),
    unit = "s") -> benchmarks

benchmarks <- data.frame(summary(benchmarks))

# Adding variable containing package name
benchmarks %>%
    mutate(package = str_extract(expr, "^[^_]+(?=_)")) -> benchmarks
```

```{r Benchmarks results}
ggplot(benchmarks, aes(x = reorder(expr, mean), y = mean, label = expr)) +
    geom_bar(stat = "identity", aes(fill=package)) +
    coord_flip() + 
    ylab("Mean time of execution (s)") + 
    xlab("Function") +
    labs(title = "Comparision of execution times of plot generations")
```



